### 2014-09-04 eval.tex
edits

\section{Evaluation}~\label{sec:eval}

{\bf Evaluation Method} We conduct our experiment with a set of large applications, e.g.,  those from the {\sf Dacapo} benchmark suite. 
The benchmarks are shown in Table. We compare our approach with the state-of-the-art technique proposed by Huang et al. As the capability of finding races depends on the monitored run, for fair comparison, we 
 apply both techniques to the same monitored run. Specifically, we use two threads in the monitored run as two threads are sufficient to manifest most of races~\cite{shanlu}.

The {\sf Dacapo} relies on the reflection functionality to execute, which imposes challenges to static analysis and transformation. We apply the tool {\sf Tamiflex}~\cite{} to overcome the problems, which resolves the reflection using the classes observed in the dynamic run. The window size is 10000.

@@

\subsection{Scalability Issues}
Our approach suffers from scalability issues for two reasons. First, as we record local access events in the monitoring run,  the monitoring run takes long time, e.g., more than 2 hours, and our trace is typically large, e.g., at the GB level. To avoid the out of memory issue during the monitoring run, we store the event to the database on the fly, rather than storing them in a buffer. Besides, we terminate the monitoring run with a shutdown hook if it is beyond 30 minutes. Second,  the prediction phase needs to load the trace to memory, which easily causes out of memory problem. To solve the problem, we separate the analysis into several runs, each run restarting the JVM and resuming from the window of last run. Different runs communicate the window id using the external file. These runs are organized together automatically using the ant. The outputs of the files are merged into one summary finally. Besides,

\begin{table}[htbp]
\caption{haha}
\begin{tabular}{|l|l|}
\hline
ad & fda \\ \hline
\multicolumn{1}{|r|}{1} & \multicolumn{1}{r|}{2} \\ \hline
\end{tabular}
\label{tb:ref}
\end{table}

\section{Case studies}
We manually inspect the reported races in small applications to gain better understanding about our approach.

@@

>>>>

\section{Evaluation}~\label{sec:eval}
Our evaluation focuses on the effectiveness  and scalability of our approach.
To measure the effectiveness, we compare with the predictive analysis, {\sf RV}~\cite{}. We chose RV for two main reasons: (1) {\sf RV} is the only open source predictive analysis, (2) {\sf RV} represents the state-of-the-art approach, which is theoretically proven to have higher detection capability than other approaches.

{\bf Evaluation Method} We conduct our experiment on a large set of applications, which are also used to evaluate {\sf RV}. Specifically, the set includes large applications such as {\sf Jigsaw}, {\sf Xalan}, {\sf Lusearch}. To handle large applications that make use of reflection, we applied {\sf Tamiflex}~\cite{}, which replaces reflective calls with the concrete method invocations recorded in the observed run for the purpose of effective static analysis.  We omit the benchmark {\sf eclipse} because the current version of {\sf Tamiflex}  leads to abnormal execution after the instrumentation, which throws an exception when the main service is started. By applying our tool to such abnormal execution, we identify only 3 races, similarly to the report of RV~\cite{}.
In addition, the {\sf montecarlo} benchmark requires an external input that we downloaded from the internet and simplified. For the large applications, we utilized the most lightweight available configuration.

 As the detection capability  depends on the observed run, for fair comparison we monitor the execution once by recording all necessary information required by both approaches, and then apply both techniques to the monitored run. Besides, our reported data for RV may be different from the original report for two reasons: (1) the different observed runs lead to different sets of races, (2) the original implementation of RV contains a bug in identification of the branches, which leads to  misses of many branches and  incorrect reduction of dependences during the analysis. We confirmed the bug with the author and fixed the bug in our experiments.\footnote{We have created an in-depth report on the bug, along with a witness test case. (See:  \url{https://sites.google.com/site/recipe3141/}.)}

@@

%Specifically, we use two threads in the monitored run as two threads are sufficient to manifest most of races~\cite{shanlu}.

\subsection{Effectiveness}
Table~\ref{tab:main} shows the main results of our analysis, which includes four sections: Trace (details about the trace), Races (detected races), Difference (comparison with RV) and Running time (the time taken by the analysis).

The Trace section includes the number of threads ($Th$), the number of shared reads ($Reads$) and shared writes ($Writes$), the subset of shared reads that read the base object references  ($Base$), where the  base object references are references used as the base/target in the following field reference or the method invocation,  the number of branches ($Br$), the synchronization events ($Sync$) and the total number of events ($Total$), which includes local accesses in addition to the aforementioned events. Specifically, for the branch, we report it in the form of $A/B$, where $A$ refers to the number of branches used by our analysis and $B$ refers to the number of branches used by RV. To ensure that the predicted run sees the same base object at each shared read/write, {\sf RV} inserts the artificial branch immediately in front of each shared access (and array accesses). We do not use such artificial branches.

We make interesting observations about the Trace section. The non-local events, i.e., all the events listed in Table~\ref{tab:main}, occupy around 30\% of the total trace in the first 7 benchmarks, but occupy less than 1\% of the trace in the rest benchmarks, which have relatively more complex logic. The reads of the shared base objects occupy a small portion ($1/3$-$1/10$) of the total shared reads. The rest shared reads read only the primitive values or 
the references that are not the base references, e.g., the references involved only in the nullness check. Besides, our analysis involves significantly less branch events as compared to the {\sf RV} approach. The difference plays an important role in the detection, which we will explain shortly.

The $Races$ section shows the number of races detected by {\sf Recipe-s}, i.e., {\sf Recipe} without exploring un-executed paths, {\sf Recipe}, i.e., the fully-fledged version, and {\sf RV}. By comparing the {\sf Recipe-s} and {\sf Recipe}, we find {\sf Recipe} finds 100+ more races, which demonstrates the strength of exploring  un-executed paths. Intuitively,  {\sf Recipe-s} predicts based on a single trace, while {\sf Recipe} predicts based on multiple traces containing different execution paths. We also compare the {\sf Recipe} version with {\sf RV} version, as illustrated in the $Difference$ section, where $Diff$ shows the races found by {\sf Recipe} but missed by {\sf RV}, $Diff'$ shows the races found by {\sf RV} but missed by {\sf Recipe}.

First of all, {\sf Recipe} finds many 150+ more races than {\sf RV}. The reasons are multifold: (1) {\sf Recipe} can reason about the accesses in the un-executed paths, while {\sf RV} and {\sf Recipe-s} can only reason about the accesses in the executed paths. (2) {\sf Recipe} or {\sf Recipe-s} allows the relaxation of the scheduling even if it breaks the inter-thread read/write dependence in the observed run. {\sf Recipe} allows the majority of the shared reads, i.e., the reads of primitive values, to freely read from a different value from a different write as long as the value leads to the same branch decisions. {\sf RV}, however, requires them to read the same values as in the observed run. (3) An critical optimization proposed by {\sf RV} is to preserve dependences only for the reads before the preceding branches  of the racy events, rather than all the reads. However, this optimization is underplayed by the fact that {\sf RV} introduces huge amount of artificial branches, i.e., one before each shared field access, which ensures the use of the same base objects.
We get rid of such artificial branches and instead, rely on the small amount of base read events to ensure the use of the same base objects (Section~\ref{}). Our strategy reduces the number of branches greatly and amplifies the effectiveness of the optimization. We also conduct case studies (Section~\ref{}) to better illustrate the scenarios.

Another interesting observation is, although {\sf Recipe} should produce all races found by {\sf RV} in theory, {\sf Recipe} may miss races found by {\sf RV} in practice (Column $Diff'$), i.e., {\sf Recipe} is not strictly more effective than {\sf RV} in practice. The underlying reason is due to the limits of the constraint solver: (1) the solver cannot compute constraints with very complex arithmatic operations (2) the solver does not support some program constants such as the scientific notation, $3E-10$.

The last section, Running time, compares the analysis time for both approaches. We find our approach is significantly slower than {\sf RV}, e.g., {\sf RV} often finishes within 200 seconds, while our approach may take more than 1 hour. This is because our approach needs to reason about the computation among variables inside the local access events, while {\sf RV} needs to only reason about the order relations among the events.

{\bf Summary of advantage and weakness\ } In general, our approach is flexible, which allows the relaxation of schedules and paths based on the fine-grained reasoning of value flows. As a result, it detects many more races compared to existing approaches. On the other hand, it is heavy weighted. The suggestion is to combine it with the lightweight approach with soundness guarantee such as {\sf RV}, by treating {\sf RV} as the preprocessing and instructing {\sf Recipe} to skip those confirmed by {\sf RV}. In this way, we can complement {\sf Recipe} by finding those missed by it due to the limit of the constraint solver.

%TODO Reorder the columns: RV ReceipeS Recipe

\begin{table*}[htbp]
\caption{Relaxed Analysis}
\begin{flushleft}
\begin{tabular}{|l|r|r|r|r|l|r|r|r|r|r|r|r|r|r|}
\hline
 & \multicolumn{ 7}{c|}{Trace} & \multicolumn{ 3}{c|}{Races} & \multicolumn{ 2}{c|}{Difference} & \multicolumn{ 2}{c|}{Running time (sec)} \\ \hline
Benchmarks & \multicolumn{1}{l|}{Th} & \multicolumn{1}{l|}{Reads} & \multicolumn{1}{l|}{Writes} & \multicolumn{1}{l|}{Base} & Br & \multicolumn{1}{l|}{sync} & \multicolumn{1}{l|}{Total} & \multicolumn{1}{l|}{Recipe-s} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} & \multicolumn{1}{l|}{Diff} & \multicolumn{1}{l|}{Diff'} & \multicolumn{1}{l|}{Recipe} & \multicolumn{1}{l|}{RV} \\ \hline
critical & 3 & 13 & 7 & 5 & 2/13 & 6 & 78 & 8 & 8 & 8 & \multicolumn{1}{c|}{0} & 0 & 8 & \multicolumn{1}{c|}{2} \\ \hline
airline & 11 & 45 & 15 & 4 & 32/82 & 30 & 317 & 9 & 9 & 9 & 0 & 0 & 490 & 4 \\ \hline
account & 3 & 46 & 21 & 10 & 3/47 & 6 & 227 & 2 & 5 & 5 & 3 & 3 & 41 & 4 \\ \hline
pingpong & 4 & 7 & 7 & 3 & 0/15 & 6 & 111 & 1 & 1 & 1 & 0 & 0 & 19 & 1 \\ \hline
bbuffer & 4 & 640 & 118 & 10 & 634/1.1K & 217 & 3.3K & 13 & 25 & 9 & 21 & 5 & 62 & 5 \\ \hline
bubblesort & 26 & 1.3k & 966 & 121 & 155/2.8K & 322 & 8.4K & 7 & 7 & 7 & 0 & 0 & 3295 & 3 \\ \hline
bufwriter & 5 & 165 & 52 & 75 & 16/130 & 44 & 525 & 4 & 10 & 2 & 8 & 0 & 63 & 9 \\ \hline
mergesort & 5 & 38 & 33 & 5 & 15/472 & 28 & 1.7K & 3 & 10 & 3 & 10 & 0 & 37 & 5 \\ \hline
raytracer & 2 & 31 & 5 & 12 & 314/8,2K & 676 & 94.5K & 4 & 6 & 4 & 2 & 0 & 47 & 2 \\ \hline
montecarlo & 2 & 5 & 86 & 2 & 1.9K/38.2K & 21.1K & 1.9M & 1 & 4 & 1 & 3 & 0 & 1 & 17 \\ \hline
moldyn & 2 & 605 & 61 & 104 & 19.6K/52.6K & 62 & 203.4K & 6 & 14 & 2 & 12 & 0 & 2842 & 1 \\ \hline
ftpserver & 28 & 684 & 299 & 71 & 4.4K/233.3K & 78.2K & 3.9M & 99 & 152 & 57 & 108 & 13 & 811 & 153 \\ \hline
jigsaw & 12 & 525 & 702 & 211 & 63.2K/467.9K & 86.7K & 5.5M & 17 & 23 & 8 & 15 & 0 & 33 & 7 \\ \hline
sunflow & 9 & \multicolumn{1}{l|}{2.1K} & \multicolumn{1}{l|}{1.3K} & 473 & 201.3K/827.0K & 50K & \multicolumn{1}{l|}{7.1M} & 38 & 78 & 20 & 69 & 11 & 4520 & 22 \\ \hline
xalan & 9 & \multicolumn{1}{l|}{1.4K} & \multicolumn{1}{l|}{0.9K} & 209 & 15.7K/103.2K & \multicolumn{1}{l|}{190.1K} & \multicolumn{1}{l|}{6.6M} & 2 & 6 & 2 & 4 & 0 & 5317 & 10 \\ \hline
lusearch & 10 & \multicolumn{1}{l|}{2.3K} & \multicolumn{1}{l|}{0.5K} & 715 & 22.2K/164K & \multicolumn{1}{l|}{93.2K} & \multicolumn{1}{l|}{9.1M} & 27 & 49 & 14 & 38 & 5 & 5430 & 8 \\ \hline
\end{tabular}
\end{flushleft}
\label{tab:main}
\end{table*}

\subsection{Case studies}
We manually inspect the reported races in small applications to gain better understanding about our approach.

@@

{\bf Relaxing the Paths\ } Figure~\ref{fig:relax2} demonstrates how our approach relaxes the schedulings to account for the unexecuted paths. 
Each thread invokes the method {\sf Sorting}, which recursively starts two children threads if there are two more available entries in the thread pool (lines 8-11), or  starts one child thread if there is only one available entry (lines 4-5).  The availability is computed through the static method {\sf available} with the constant {\tt total}, which is equal to 5. The shared variable {\tt alive} denotes the used entries.

Initially there is one sorting thread. After it starts Thread 1 and Thread 2, there are three threads alive and only two more entries are available.
Suppose in the observation run, Thread 2 consumes both thread entries and starts the children Thread 3 and Thread 4 (not shown), updating {\tt alive} to 5, then Thread 1 cannot execute the branch at lines 4-5. Huang et al.~\cite{} require the predicted run to preserve the dependence denoted by the dotted line since the dependence affects the branch condition at line 2. As a result, the predicted run follows the same branch decision and cannot reason about the code at lines 4-5. Our analysis does not have such limitation. Instead, it allows the predicted run to reason about the un-explored code. Specifically, it does not need to preserve the dependence from line 11 to line 1 and it allows line 1 (Thread 1) to read from line 9 (Thread 2). As a result, the branch condition guarding the unexplored branch is evaluated to be true, enabling the unexplored path in the constraint solver. Finally, the solver identifies the race between line 5 (Thread 1) and line 1 (Thread 3). Note that the two lines are synchronized on different locks~\footnote{We abbreviate the {\tt synchronized} keyword as {\tt sync}.
}.

\begin{figure}[htp]
\centering
\includegraphics[scale=0.7]{cases/Visio-msort.pdf}
\caption{Relaxation of Paths}\label{fig:relax2}
\end{figure}

### 2014-09-04 guarantee.tex
edits

\section{Theoretical Guarantees}~\label{sec:guarantee}
In this section, we first prove the soundness of \tool and then discuss the detection capability as compared to existing approaches.
Both the proof and discussion are in the context of sequential consistency model~\cite{scmodel}.

>>>>

\section{Theoretical Guarantees}~\label{sec:guarantee}
In this section, we first prove the soundness of \tool\ and then discuss the detection capability as compared to existing approaches.
Both the proof and discussion are in the context of sequential consistency model~\cite{scmodel}.

### 2014-09-04 relax1.tex
edits

\section{Relaxation of Flow Dependencies}~\label{sec:relax1}
The main goal of our analysis is to, given an original trace $\tau$, derive a trace $\tau'$ over the same set of events in $\tau$, which has a new scheduling  of the events (the order of events inside each thread should remain unchanged) and a new mapping from variables to values in the events. Our first relaxation comes from the insight that, we allow the variables to read different values than in the original trace even if such variables are used to determine the control flow, while existing approaches enforce them to read the same values as in the original trace. As a result, our relaxation allows more schedules, which are likely to expose more races.

Not every scheduling or every mapping  leads to a new feasible trace $\tau'$, therefore, we need to compute the scheduling and mapping that lead to feasible trace. We explain the technique details in the following.

@@

\subsection{SSA Form of the Trace}~\label{sec:ssa}
First of all, we require the SSA form of the trace, i.e., each variable in defined exactly once in the trace. Such requirement is a prerequisite for the computation of the mapping, in which each variable is mapped to exactly one value. Another side effect is that the def/use chains become explicit in the SSA form, which simplifies the following analysis steps.

\tool\ handles the local assignment and heap accesses differently.
\begin{itemize}
\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the loops and recursions are fully resolved in a concrete trace, obviating the need for the {\sf Phi} node. More concretely, we replace the variable $v$ defined in an event, as well as the following uses of the definition, to a new variable $v^{id}$, where $id$ is the unique id of the event. The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once. Note that the uses of a definition can be computed easily by scanning the trace afterwards for the variables of the same name before the next redefinition. 
\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations behave differently, therefore, we introduce different SSA form for them. 
\begin{itemize}
\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to the single thread. Therefore we model them as local assignments: We introduce a fresh local variable {\tt $l_{o.f }$} to replace each definition and  the corresponding uses of the location of $o.f$ ($o$ denotes an object referenced by the variable $x$ in $x.f$).  
%An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification %(Section~\ref{sec:relax1}).
\item  {\bf Shared Heap Accesses\ } The accesses of shared locations are more complex. Each shared read may read from one of multiple writes that update the shared location, under different schedules.     Given the write $x.f=y$ or the read $y=x.f$, we introduce two symbols, $W^{id}_{\tt o.f }$, which denotes the write access by the event $id$ to the  shared heap location  $o.f$  ($o$ denotes an object referenced by the variable $x$ in $x.f$), and $R^{id}_{\tt  o.f }$,  which denotes the read access  by the event $id$ to the location $o.f$. 
\end{itemize}
We use the heap location such as $o.f$ in the the symbolic form of heap accesses. An underlying assumption is that, the heap location remains unchanged in our predictive analysis, which we ensure through additional constraints (Section~\ref{sec:relax1}).
\end{itemize}

@@

Figure~\ref{fig:t4running2} exemplifies a trace and its SSA form, where the trace is generated from the program in Figure~\ref{fig:running2}.
Each label in  Figure~\ref{fig:t4running2} on the left hand denotes the id of the event. As seen, event $e_5$ uses the variable $s^2$ defined at event 2 and defines the variable $s^5$, which is used at $e_8$. The read/write of the shared variables are denoted in the special form, such as $W^{8}_y$ and  $R^{9}_y$.

\subsection{Constraint System}~\label{sec:constraints}
Based on the SSA form of the trace, we build the constraints to compute a new trace with the new schedule and new mapping of variables to values.
To model the schedule, we introduce the order variable $O_{id}$ for each event $e_{id}$. Given two events $e_i, e_j$ from two different threads, $O_{e_i}<O_{e_j}$ in a trace means $e_i$ is scheduled before $e_j$. We omit the synchronization constraints purposely, as they are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.

@@

{\bf Race Condition\ } Following the definition, a pair, ($e_i$, $e_j$), forms a race iff (1) $e_i$ and $e_j$ are accesses of the same location $\ell$ by different threads, (2) at least one of them writes to $\ell$, and (3) $e_i$ and $e_j$ run concurrently. We refer to the candidate pair throughout this section and call the two events as the racy events.

We first identify all candidate pairs of events that access the same location from different threads (and at least one is a write), then we check each pair  separately.   The checking is achieved by encoding all necessary constraints and invoking a constraint solver.  The first constraint asserts the feasibility of the concurrent execution of the racy events. 
$$
O_{i} = O_{j}
$$

{\bf Intra-thread Constraints\ } Given the candidate race pair, $e_i$ and $e_j$, suppose $e_i$ is after $e_j$ in the original trace $\tau$, the predictive analysis reschedules the events in the prefix of $e_i$, i.e., prior to $e_i$, so that $e_i$ and $e_j$ can run concurrently. The predictive analysis by design requires that each thread should follow the same event sequence prior to $e$ as in the original run.  The preservation of the event sequence for each thread requires the following intra-thread constraints:

@@

More formally,  we require that 
$$
\bigwedge_{e_k \in \tau\downarrow_{\leq i} \wedge  type^{e_k}=branch.\ }  inst^{e_k} 	\equiv \lsyn inst \rsyn^{e_k}
$$ 
where  the branch event $e_k$ is used as a boolean expression. For example, given the branch {\tt if(x<y)\dots}, the constraint is in the form of $(x<y)\equiv true$, suppose the branch is evaluated as true in the original trace. The constraint specifies that the boolean expression in the predicted trace should be evaluated to the same boolean value as in the original trace. Importantly, different from existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but adopt the relaxed requirement that the evaluation of branching expressions remains the same. For example, suppose  the branch event $x<y$ takes the value $1<2$ in the original trace, existing analyses require the same values for the variables $x$ and $y$ in the predicted run, while we allow  the branch event to take other values of the variables, such as $3<4$. 
\item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that
$$
\begin{array}{rl}
\forall e_m, e_n \in \tau\downarrow_{\leq i}, s.t., & t^{e_m} = t^{e_n}. \\
& m < n}\  \Rightarrow O_{m} < O_{n}
\end{array}
$$ 
The constraint specifies that the two events from the same thread should follow the same order as reflected by the ids of the events. Note that $m$ in $e_m$ denotes its id. Again, we only consider the events in the prefix of $e_i$.
\item {\bf Intra-thread Data Flow Constraints\ } The mapping of local variables should not contradict the data flow semantics of each instruction.   More formally, we require that
$$
\bigwedge_{e_k \in \tau\downarrow_{\leq i} \wedge type(e_k)=assign|heapr|heapw}\
	inst^{e_k}
$$
where $inst^{e_k}$, such as $x=y+z$ or $R^{id}_{x}=y$, captures the constraint over the values of the variables imposed by the instruction. For example, $x=y+z$ is not satisfied if the solver maps the three variables to $3, 4, 5$ respectively. Again we consider only the events in the prefix of  $e_i$. Specially, for the instruction involving object creation, $x=new (...)$, we encode the object as a unique integer that denotes its heap address, calculated by $System.identityHashcode(x)$ at runtime.
\end{itemize}

%Rescheduling through Relaxation of Flow Dependencies
{\bf Inter-thread Data Flow Constraints for Relaxation\ } We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace. More precisely, \tool\ is able to relax value flow dependencies in the original trace:  a read access may read a
different value from other write events, as long as the read value enforces feasibility.
 This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access must read the same  value (often from the same write event) as in the original trace.

To ensure feasibility under relaxation of flow dependencies, we need to secure the flow between the read/write with the execution schedule. 
For example, in Figure \ref{fig:running}, for the read access to ${\tt y}$ at $e_4$ obtains the value assigned to ${\tt y}$ at $e_1$, we need the schedule where 
$e_4$ happens after $e_1$ and other writes such as $e_3$ do not interleave them ($e_3$ can only happen after $e_4$ in this case).

@@

An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning, i.e., they involve different base objects and do no longer match with each other. As an illustration, we refer to Figure \ref{fig:heapAccess}. While the read at the event $e_7$ appears to match the write at $e_5$, this is conditioned on the read at $e_6$ being linked to the assignment at $e_4$. Suppose  the event $e_6$ reads from $e_3$ due to the relaxation, then $e_7$ and $e_5$ no longer share the same base object (or the same location). Even worse, we do not know what events $e_7$ match because the base object becomes unknown/symbolic.

@@

To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in the field reference before the racy events retain their original address in the predicted trace.
This achieves two guarantees: First, matching heap access events in the original trace are guaranteed to also match in the predicted trace. Second, candidate races in the original trace remain viable in the predicted trace as they still refer to same location. Third, the sharedness nature of each heap access remains the same because
each location is accessed by the same number of threads given that the base object is unchanged.

>>>>

\section{Relaxation of Flow Dependencies}~\label{sec:relax1}
The main goal of our analysis is to derive, given an original trace $\tau$, a trace $\tau'$ over the same set of events in $\tau$, which has a new scheduling  of the events (the order of events inside each thread should remain unchanged) and a new mapping from variables to values in the events. Our first relaxation comes from the insight that we allow the variables to read different values than in the original trace even if such variables are used to determine the control flow, while existing approaches enforce them to read the same values as in the original trace. As a result, our relaxation allows more schedules, which are likely to expose more races.

Not every scheduling or every mapping  leads to a new feasible trace $\tau'$. Therefore, we need to compute scheduling and mapping instances that lead to feasible traces. We achieve this via constraint solving. We explain the details of our technique in the following.

@@

\subsection{SSA Form of the Trace}~\label{sec:ssa}
First of all, we rewrite the trace into SSA form, such that every variable in defined exactly once in the trace. This requirement is a prerequisite for the computation of the mapping, in which each variable is mapped to exactly one value. Another side effect is that the def/use chains become explicit in the SSA form, which simplifies the following analysis steps.

\tool\ handles the local assignment and heap accesses differently.
\begin{itemize}
\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the loops and recursions are fully resolved in a concrete trace, obviating the need for the {\sf Phi} node. More concretely, we replace the variable $v$ defined in an event, as well as the following uses of the definition, to a new variable $v^{id}$, where $id$ is the unique id of the event. The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once. Note that uses of a definition can be computed easily by scanning the trace suffix for accesses to the same variable name before its next definition. 
\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations play different roles in our analysis. We therefore introduce different SSA form for them. 
\begin{itemize}
\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to a single thread. Therefore we model them as local assignments: We introduce a fresh local variable {\tt $l_{o.f }$} to replace each definition and  the corresponding uses of the location of $o.f$ ($o$ denotes an object referenced by the variable $x$ in $x.f$).  
%An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification %(Section~\ref{sec:relax1}).
\item  {\bf Shared Heap Accesses\ } The accesses of shared locations are more complex. Each shared read may read from one of multiple writes that update the shared location under different schedules (e.g., the read access to {\tt y} at line 4 in Figure \ref{fig:running}, which can either obtain the definition before line 1 or at line 1 or at line 3).   Given the write $x.f=y$ or the read $y=x.f$, we introduce two symbols, $W^{id}_{\tt o.f }$, which denotes the write access by the event $id$ to the  shared heap location  $o.f$  ($o$ denotes an object referenced by the variable $x$ in $x.f$), and $R^{id}_{\tt  o.f }$,  which denotes the read access  by the event $id$ to the location $o.f$. 
\end{itemize}
We use the heap location such as $o.f$ in the the symbolic form of heap accesses. An underlying assumption is that the heap location remains unchanged in our predictive analysis, which we ensure through additional constraints (Section~\ref{sec:relax1}).
\end{itemize}

@@

Figure~\ref{fig:t4running2} exemplifies a trace and its SSA form, where the trace is generated from the program in Figure~\ref{fig:running2}.
Each label in  Figure~\ref{fig:t4running2} on the left hand denotes the id of the event. As seen, event $e_5$ uses the variable $s^2$ defined at event $e_2$ and defines the variable $s^5$, which is used at $e_8$. Reads and writes of shared variables are denoted in the special form explained above (e.g.: $W^{8}_y$ and  $R^{9}_y$).

\subsection{Constraint System}~\label{sec:constraints}
Based on the SSA form of the trace, we build the constraints to compute a new trace with the new schedule and new mapping of variables to values.
To model the schedule, we introduce the order variable $O_{id}$ for each event $e_{id}$. Given two events $e_i, e_j$ from two different threads, $O_{e_i}<O_{e_j}$ means that $e_i$ is scheduled before $e_j$. We omit the synchronization constraints intentionally, as they are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.

@@

{\bf Race Condition\ } Following the standard definition, pair ($e_i$, $e_j$) of events forms a race iff (1) $e_i$ and $e_j$ are accesses of the same location $\ell$ by different threads, (2) at least one of them writes to $\ell$, and (3) $e_i$ and $e_j$ run concurrently. We refer to the candidate pair throughout this section as (candidate) \emph{racy events}.

We first identify all candidate pairs of events that access the same location from different threads (and at least one is a write). We then check each pair  separately.   The checking is achieved by encoding all necessary constraints and invoking a constraint solver.  The first constraint asserts the feasibility of the concurrent execution of the racy events:
$$
O_{i} = O_{j}
$$

{\bf Intra-thread Constraints\ } Given the candidate race pair, $e_i$ and $e_j$, suppose $e_i$ is after $e_j$ in the original trace $\tau$. The predictive analysis reschedules the events in the prefix of $e_i$, i.e., prior to $e_i$, so that $e_i$ and $e_j$ can run concurrently. The predictive analysis by design requires that each thread should follow the same event sequence prior to $e$ as in the original run.  The preservation of the event sequence for each thread requires the following intra-thread constraints:

@@

More formally,  we require that 
$$
\bigwedge_{e_k \in \tau\downarrow_{\leq i} \wedge  type^{e_k}=branch.\ }  inst^{e_k} 	\equiv \lsyn inst \rsyn^{e_k}
$$ 
where  the branch event $e_k$ is used as a boolean expression. For example, given the branch {\tt if(x<y)\dots}, the constraint is in the form of $(x<y)\equiv {\bf true}$ assuming the branch evaluates to {\bf true} in the original trace. The constraint specifies that the boolean expression in the predicted trace should be evaluated to the same boolean value as in the original trace. Importantly, Unlike existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but adopt the relaxed requirement that the evaluation of branching expressions remains the same. For example, suppose  the branch event $x<y$ takes the value $1<2$ in the original trace. Existing analyses require the same values for the variables $x$ and $y$ in the predicted run, while we allow  the variables flowing into the branch event to assume other values, such as $3<4$, as long as the branch expression retains the same truth value. 
\item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that
$$
\begin{array}{rl}
\forall e_m, e_n \in \tau\downarrow_{\leq i}, s.t., & t^{e_m} = t^{e_n}. \\
& m < n\  \Rightarrow O_{m} < O_{n}
\end{array}
$$ 
This constraint specifies that two events from the same thread should follow the same order as reflected by the ids of the events. Note that $m$ in $e_m$ denotes its id. Again, we only consider the events in the prefix of $e_i$.
\item {\bf Intra-thread Data Flow Constraints\ } The mapping of local variables should not contradict the data flow semantics of each instruction.   More formally, we require that
$$
\bigwedge_{e_k \in \tau\downarrow_{\leq i} \wedge type(e_k)=assign|heapr|heapw}\
	inst^{e_k}
$$
where $inst^{e_k}$, such as $x=y+z$ or $R^{id}_{x}=y$, captures the constraint over the values of the variables imposed by the instruction. For example, $x=y+z$ is not satisfied if the solver maps the three variables to $3, 4, 5$ respectively. Again we consider only the events in the prefix of  $e_i$. Specially, for the instruction involving object creation, $x=new (...)$, we encode the object as a unique integer that denotes its heap address, calculated by $System.identityHashcode(x)$ at runtime.
\end{itemize}

%Rescheduling through Relaxation of Flow Dependencies
{\bf Inter-thread Data Flow Constraints for Relaxation\ } We now move to the first novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace. More precisely, \tool\ is able to relax value flow dependencies in the original trace:  a read access may read a
different value from other write events, as long as the read value enforces feasibility.
 This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access must read the same  value (often from the same write event) as in the original trace.

To ensure feasibility under relaxation of flow dependencies, we need to secure the flow between the read/write with the execution schedule. 
For example, in Figure \ref{fig:running} for the read access to ${\tt y}$ at $e_4$ to obtain the value assigned to ${\tt y}$ at $e_1$, we need a schedule where 
$e_4$ happens after $e_1$ and other writes such as $e_3$ do not interleave them. ($e_3$ can only happen after $e_4$ in this case.)

@@

An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning, i.e., they involve different base objects and no longer match with each other. As an illustration, we refer to Figure \ref{fig:heapAccess}. While the read at the event $e_7$ appears to match the write at $e_5$, this is conditioned on the read at $e_6$ being linked to the assignment at $e_4$. Suppose  the event $e_6$ reads from $e_3$ due to the relaxation. Then $e_7$ and $e_5$ no longer share the same base object (or the same location). Even worse, we do not know which events $e_7$ matches, because the base object is no longer known.

@@

To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in field access statements before the racy events retain their original address in the predicted trace.
This achieves two guarantees: First, matching heap access events in the original trace are guaranteed to also match in the predicted trace. Second, candidate races in the original trace remain viable in the predicted trace as they still refer to same location. Third, sharing between threads of heap locations remains unchanged  because
each location is accessed by the same number of threads given that the base object is the same.
